Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-25 paper/high-assurance-contracts.tex
finalized wellformedness discussion

\begin{definition}[Code contract correctness]
Assume code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$.
The contract is \emph{correct}, if for all $E$ and $t$, whenever the
assumptions (historically) hold in $E$ and evaluation steps from $E$
to $E'$, then the guarantees hold in $E'$:
\[
\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
\]
\end{definition}
\footnote
{KLS: At this point we would like to assert that the AGREE ``consistency
 check'' for a contract implies code contract correctness.}

How does this notion of correctness integrate with the verification
conditions addressed by model checking? In \emph{contract
correctness}, the model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intuition, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program $(\mathit{Eqs}
\cdot G)$ from a code contract we find ourselves in the tepid
bathwater of imperative program verification. For example, at this
point one could apply a verification condition generator to help break
down the proof obligation. In order to use the code contract result in
system contract verifcation conditions, one has to show that the code
contract result implies the system verification condtion.

\subsubsection*{Wellformedness}
The evaluation semantics given above is based on the stream
computation model, where values arbitrarily `deep in the past' can be
accessed in computations by use of $\konst{pre}(-)$. However, the
conventional program languages we target do not support such a
view. We are currently examining transformations that compile such
deep temporal accesses away by introducing intermediate
variables. Avoiding discussion of the algorithm, we focus on the
notion of \emph{wellformed} equations that needs to hold. The guiding
intuition is that an \emph{implementable} {\Eqs} can allow access at
most one step in the past, and accesses of a past value for a variable
are only possible until the variable is updated. In lieu of a formal
definition, we give a few examples.

\begin{definition}[Wellformedness]

\end{definition}

>>>>

\begin{definition}[Code contract correctness]
Assume code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$.
The contract is \emph{correct}, if for all $E$ and $t$, whenever the
assumptions (historically) hold in $E$ and evaluation steps from $E$
to $E'$, then the guarantees hold in $E'$:
\[
\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
\]
\end{definition}

How does this notion of correctness integrate with the verification
conditions generated by AGREE? In \emph{contract correctness}, the
model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intuition, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program $(\mathit{Eqs}
\cdot G)$ from a code contract we find ourselves in the tepid
bathwater of imperative program verification. For example, at this
point one could apply a verification condition generator to help break
down the proof obligation. In order to use the code contract result in
system contract verifcation conditions, one has to show that the code
contract result implies the system verification condtion.

\subsubsection*{Wellformedness}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages we target do not directly support streams.  We are currently
developing a means of translating away such deep temporal accesses by
the introduction of intermediate variables. For now, we focus on the
notion of \emph{wellformed} equations that needs to hold. The guiding
intuition is that a wellformed {\Eqs} allows access at most one step
in the past, and accesses of a past value for a variable are only
possible until the variable is updated. In informal terms, this means
that ${\Eqs}\cdot G$ in a code contract must be such that:
\begin{itemize}
\item \konst{fby} only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed before the defining equation $v = e$, and may occur in $e$.
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-wellformed manner, by
{\small
\begin{verbatim}
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{verbatim}
}
\noindent or by (again non-wellformed)
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib)).
\end{verbatim}
}
\noindent or by (wellformed):
{\small
\begin{verbatim}
  N = 0 -> 1 + pre(N)
  A = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(A)
\end{verbatim}
}
A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.
